---
title: "stats_ws_seasons_species"
author: "FA"
date: "8 4 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      results = 'show',
                      message = FALSE,
                      warning = FALSE ) # this sets settings for all the script

library(lme4)
library(nlme)
library(dplyr)
library(ggplot2)
library(pander)
library(car)
library(lmerTest)
library(MuMIn)


```



```{r data, echo=FALSE}
atpu = readRDS("ATPU_bird_data_ringyear_season.rds")
blki = readRDS("BLKI_bird_data_ringyear_season.rds")
# one value per ringyear and per season 

atpu$species = rep("ATPU", nrow(atpu))
blki$species = rep("BLKI", nrow(blki))

dat = rbind(atpu,blki)
rm(atpu,blki)


dat = dat %>% dplyr::filter(!is.na(ws))

dat$colony_species = paste(dat$colony,dat$species,sep = "_")

sub = dat %>% dplyr::select(colony_species,year2, ringyear, ws, season, species)
print(head(sub))



```

## Data exploration

Cf Zuur et al 2010 MEE

```{r exploration}
# 1. outliers
dotchart(dat$ws)

# 2. homogeneity of variance

ggplot(dat, aes(x=season, y = ws)) + geom_boxplot() + facet_grid(cols=vars(species))

# 3. normality 

ggplot(dat, aes(x=ws)) + geom_histogram(binwidth = 0.2)

table(dat$season,dat$species)

ggplot(dat, aes(x=ws)) + geom_histogram() + facet_grid(rows = vars(species,season))



```



## Choice of the random effect structure:

We include three random variables: year, colony and invididual, with individual nested in colony.



```{r random-effect}
modGLS = gls(ws ~ 1 + season * species, data = dat, method = "REML")
AIC(modGLS)

modA = lmer(ws ~ season * species + (1 | colony/ring) + (1|year2), data = dat,REML = T)
AIC(modA)


```

## Get fixed effects structure


```{r fixed-effect}
mod0 = lmer(ws ~ 1 + (1 | colony/ring) + (1|year2), data = dat,REML = F)
AIC(mod0)

mod1 = lmer(ws ~ season * species + (1 | colony/ring) + (1|year2), data = dat,REML = F)
AIC(mod1)

mod2 = lmer(ws ~ season + species + (1 | colony/ring) + (1|year2), data = dat,REML = F)
AIC(mod2)

mod3 = lmer(ws ~ season  + (1 | colony/ring) + (1|year2), data = dat,REML = F)
AIC(mod3)

mod4 = lmer(ws ~ species + (1 | colony/ring) + (1|year2), data = dat,REML = F)
AIC(mod4)


```

## best model in REML

```{r best}

mod1 = lmer(ws ~ season * species + (1 | colony/ring) + (1|year2), data = dat,REML = T)
AIC(mod1)
summary(mod1)

confint(mod1, oldNames=FALSE)

plot(mod1)
qqnorm(resid(mod1))
qqline(resid(mod1)) # not great but ok 


hist(resid(mod1),breaks=20)


dat$resid = resid(mod1)

boxplot(dat$resid~dat$colony)
ggplot(dat, aes(x=season, y = resid)) + geom_boxplot() + facet_grid(cols = vars(species))
ggplot(dat, aes(x=year2, y = resid)) + geom_boxplot() 

vif(mod1)

r.squaredGLMM(mod1)

```


## graph with mean wind support and CI per species and season

Mean +/- CI per species and per season

```{r graph, echo=FALSE, eval = FALSE}
meanCI = function(x,myquantile){
  bootmed = apply(matrix(sample(x, rep=TRUE, 10^4*length(x)), nrow=10^4), 1, mean, na.rm = T)
  unname(quantile(bootmed, myquantile, na.rm = T))
  
}


dat_spec_seas = dat %>%  
  group_by(season,species) %>% 
  summarise(
    ws.CI025 = meanCI(ws,0.025),
    ws.CI975 = meanCI(ws,0.975),
    sd.ws = sd(ws,na.rm=T),
    ws = mean(ws,na.rm=T),
    n = n()) 

dat_spec_seas$species2 = ifelse(dat_spec_seas$species=="ATPU", "b. Puffin", "a. Kittiwake")
dat_spec_seas$se = dat_spec_seas$sd.ws/sqrt(dat_spec_seas$n)



```


## violin plot with mean wind support and CI per species and season

Violin plot asked by reviewer
Mean +/- CI per species and per season

```{r graph 2, echo=FALSE}

dat$species2 = ifelse(dat$species=="ATPU", "Puffin", "Kittiwake")

# trial with another piece of code https://dbaranger.medium.com/showing-your-data-scatter-box-violin-plots-1f3bb06c8c2b


myplot = ggplot(data = dat,aes(x= season, y = ws, fill = species2)) +
     scale_fill_manual( values = c("yellow","red"), name ="") +
     geom_violin(alpha=0.4, position = position_dodge(width = .85),color="darkgrey", size = 0.2) +
     geom_boxplot( width = 0.2,  outlier.size = 0.1, color="black", position = position_dodge(width = .85), notch = TRUE, size = 0.5)  +
  geom_hline(yintercept =0, linetype="dashed") +
  labs(x = "Migratory season", y=expression(Wind~support~(m.s^{-1}))) +
   theme_bw() +
  theme( panel.grid.major = element_blank(),panel.grid.minor = element_blank(),strip.background = element_blank())



ggsave(file = "Fig_model1_v3.jpg", plot = myplot , dpi = 300 ,width = 4, height = 3, units = "in")


```



## summary table per species

```{r table per species,eval = FALSE,echo = FALSE}

dat_spec = dat %>%  
  group_by(species) %>% 
  summarise(
    ws.CI025 = meanCI(ws,0.025),
    ws.CI975 = meanCI(ws,0.975),
    sd.ws = sd(ws,na.rm=T),
    ws = mean(ws,na.rm=T),
    n = n()) 

print(dat_spec)
```

